### 2018-09-07 README.Rmd
add README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations

This is the repository for the paper: "Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations" - you can find all the data and code to replicate the paper here.