### 2020-05-14 cover-letter/cover-letter.Rmd
update cover letter

---
from: Nicholas Tierney
fromqualifications: BPsySc (Hons), PhD (Statistics)
fromtitle: Lecturer
fromWWW: njtierney.com
fromemail: nicholas.tierney@gmail.com
fromphone: +61 3 9905 9352
fromsignature: sigfile.png
toname: Editor in Chief
toaddress: Journal of Statistical Software
opening: Dear Editors-in-Chief
closing: Regards, \newline Dr. Nicholas Tierney \newline Lecturer
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider my article "Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations" for possible publication in the Journal of Statistical Software. This article describes new work on handling missing values from a data centric approach. The methodology is implemented in the R package, `naniar`, which is available on CRAN, https://cran.r-project.org/web/packages/naniar, with the development version on github https://github.com/njtierney/naniar. `naniar` is now downloaded 14 thousand times a month, and has been downloaded over 160 thousand times. `naniar` has also seen wide praside, for example: [Karandeep Singh](https://twitter.com/kdpsinghlab/status/1110529278830170116):

> Looking forward to teaching about missing data and mechanisms of missingness today. Prep work involved swapping out the mi and VIM packages from last year's #rstats code with @nj_tierney's excellent naniar package.

And [Omar Wasow](https://twitter.com/owasow/status/1006743170846109697):

> naniar also tipped the scale for me w/teaching. In past, I've skipped lecturing on missing data. This year, I revised lectures & covered naniar/simputation/Amelia. Was grateful to easily show basic issue as well as trade-offs of different forms of imputation (eg mean vs lm). Thx!

In addition, the R package has been cited 9 times, and the pre-print of this paper has been cited 6 times since 2018.

Along with the PDF is an accompanying Rmarkdown file, which contains all of the code to reproduce the results in this paper. This article is not under consideration elsewhere.  This paper was submitted two years aho to the Journal of Computational and Graphical Statistics, but was rejected. It has now been substantially re-written. We have submitted a pre-print of this article to arxiv, and to our internal department pre-print at Monash University.

We hope that you find this paper of interest to the readers of Journal of Statistiacl Software,