### 2020-08-18 paper/review/response-to-review.Rmd
add reviewer comments

---
title: "Response to Review"
---

JSS 4108:

Nicholas John Tierney, Dianne Cook

Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations

---------------------------------------------------------
For further instructions on JSS style requirements please see the Style Files (in particular section 2.1 Style Checklist of jss.pdf) and FAQ at http://www.jstatsoft.org/about/submissions.

For further examples please see RECENT JSS papers.
---------------------------------------------------------

From the editorial team:

We apologize for the additional delays in processing 
your submission due to the COVID-19 pandemic.

The submission "Expanding tidy data principles to facilitate missing data
exploration, visualization and assessment of imputations" presents the R 
package naniar which provides data structures, summaries and visualizations 
for missing data.

Additional comments:

Please add some discussion on what kind of functionality is available in 
other statistical software packages than R in order to better embed the 
contribution not only within R, but the statistical software ecosystem in 
general.

Please provide a plain R script which when run reproduces all results 
presented in the manuscript. In general purling the Rmd file and editing 
the resulting R file  slightly in order to guide the user what each block 
does, removing knitr specific content,   etc. is suitable.

* Please add some discussion on what kind of functionality is available in 
other statistical software packages than R in order to better embed the 
contribution not only within R, but the statistical software ecosystem in 
general.

* Please provide a plain R script which when run reproduces all results 
presented in the manuscript. In general purling the Rmd file and editing 
the resulting R file slightly in order to guide the user what each block 
does, removing knitr specific content, etc. is suitable.

* The JSS style (https://www.jstatsoft.org/pages/view/style) should be 
followed more closely, especially using the title vs. sentence case, 
JSS-specific markup (\proglang, \pkg), and improving the BibTeX file. Some 
aspects of the style are handled incorrectly by "rticles", though (see
https://github.com/rstudio/rticles/issues/254).

* Note that the affiliation of the second author is misspelled.

Manuscript style comments:

o For the code layout in R publications, we typically distinguish input/output using Sinput/Soutput (or equivalently CodeInput/CodeOutput). Unless there are special reasons to format it differently, the input should use the text width (up to 76 or 77 characters) and be indented by two spaces, e.g.,

begin{Sinput}
R> example_model <- lm(response ~ variable1 + variable2 + variable3, 
+    weights = w, data = mydata)
\end{Sinput}

o If using "e.g." and "i.e." add a comma after the period to keep LaTeX from interpreting them as the end of a sentence, i.e.: "e.g., " and "i.e., ".

o The rule for capitalizing the starting letters of Figure, Section and Table is as follows: If you are referring to a particular figure/section/table then capitalize the first letter, otherwise use a lower-case first letter. For example, something shown in Section 4 vs. there are three sections in this paper.

o All captions should appear below the corresponding figure/table. The captions should be in sentence style and end with a period.  No additional formatting (such as \emph, f or \it) should be used for the caption.

o All table row/column headers should also be in sentence style. There should not be further footnote-style annotations in tables; these should all be placed in the caption.

o As a reminder, please make sure that:
	- \proglang, \pkg and \code have been used for highlighting throughout the paper (including titles and references), except where explicitly escaped.

References:

o Journal of the Royal Statistical Society B (not: Journal of the Royal Statistical Society, Series B)
	the EM Algorithm.” Journal of the Royal Statistical Society. Series B
the EM Algorithm.” Journal of the Royal Statistical Society. Series B
o Springer-Verlag (not: Springer)
	and GGobi. 1st edition. Springer Publishing Company, Incorporated.

o Please make sure that all software packages are \cite{}'d properly.
o All references should be in title style.
o See FAQ for specific reference instructions.

Code:

o Please make sure that the files needed to replicate all code/examples within the manuscript are included in a standalone replication script.