### 2021-05-19 README.Rmd
update renv parts

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations

This is the repository for the paper: "Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations" - you can find all the data and code to replicate the paper here.

- Note that you need to have the development version of `knitr` and `rmarkdown`
- Packages used described in `renv.lock` file
- you can install the packages used in the lockfile by running:

```r
library(renv)
restore()
```

This will install the packgages used in the lockfile, for this project. It should not impact your other local R installations.