### 2022-12-01 jss-copyedit/2022-01-12-4108/paper.tex
add files that have been copyedited by JSS

\documentclass[article]{jss}

\usepackage{thumbpdf,lmodern}
\graphicspath{{Figures/}} % folder with figures

\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
Nicholas Tierney\\Monash University, \\Telethon Kids Institute
\And Dianne Cook\\Monash University
}
\title{Expanding Tidy Data Principles to Facilitate Missing Data Exploration,
Visualization and Assessment of Imputations}

\Plainauthor{Nicholas Tierney, Dianne Cook}
\Plaintitle{Expanding Tidy Data Principles to Facilitate Missing Data Exploration,
Visualization and Assessment of Imputations}
\Shorttitle{Explore Missingness with \pkg{naniar}}

\Abstract{
Despite the large body of research on missing value distributions and
imputation, there is comparatively little literature with a focus on how to
make it easy to handle, explore, and impute missing values in data.  This
paper addresses this gap.  The new methodology builds upon tidy data
principles, with the goal of integrating missing value handling as a key
part of data analysis workflows.  We define a new data structure, and a
suite of new operations.  Together, these provide a connected framework for
handling, exploring, and imputing missing values.  These methods are
available in the \proglang{R}~package \pkg{naniar}. \citep{pkg:naniar}, available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=naniar}.
}

\Keywords{statistical computing, statistical graphics, data science,
data visualization, \pkg{tidyverse}, data pipeline, \proglang{R}}
\Plainkeywords{statistical computing, statistical graphics, data science,
data visualization, tidyverse, data pipeline, R}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
  Nicholas Tierney\\
  Monash University\\
  Melbourne, Victoria, Australia\\
  \emph{and}\\
  Telethon Kids Institute\\
  Perth, Western Australia, Australia\\
  nicholas.tierney\@gmail.com

  Dianne Cook\\
  Monash University\\
  Melbourne, Victoria, Australia\\
  }

% Pandoc citation processing

% Pandoc header
\usepackage{amsmath}
\usepackage[english]{babel}

\usepackage{float}
\usepackage{booktabs}

\begin{document}

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

As data science has become a more solid field, theories and principles have
developed to describe best practices.  One such idea is ``tidy data'', which
defines a clean, analysis-ready format that informs workflows converting raw
data through a data analysis pipeline \citep{Wickham2014}.  Another idea is
the grammar of graphics, describing how to map data values into a
visualization \citep{Wilkinson2012}.  These principles have been widely
adopted, but do not address the problem of missing data.  In particular,
there is little guidance on how to handle missing values in a data analysis
workflow.  Most analysis and visualization software simply drop missing
values when making a plot, although some (\pkg{ggplot2}) provide a warning
(Figure~\ref{fig:warning}).

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\linewidth]{warning-1}
\caption[How \pkg{ggplot2} behaves when displaying missing values]{How
\pkg{ggplot2} behaves when displaying missing values.  A warning message is
displayed, but missing values are not shown in the plot.}\label{fig:warning}
\end{figure}

The imputation literature focuses on ensuring valid statistical inference is
made from incomplete data.  This is approached chiefly through probabilistic
modelling, assuming the mechanism of missing data is known to the analyst.
It does not address how to explore, understand, and handle missing data
structures and mechanisms.

However, something must be known about the missing value structure to
produce a complete dataset for analysis, whether by case- or
variable-deletion, or with imputation.  Decision tree models or latent group
analysis can reveal structures or patterns of missingness
\citep{Tierney2015, Barnett2017}, but are not definitive.  While there are
times where the missing data mechanism is obvious, determining this is
typically not straightforward.  A straightforward case is where missing
values are encoded in the data in a non standard format that still indicates
missingness, such as ``N.A.'', ``N/A'', or ``missing''.  However,
identifying these amongst all data can be problematic as it is hard to know
what you are looking for, like finding a needle in a haystack.  A more
challenging case is where missing values arise due to the increasing or
decreasing values of other variables (including themselves).  For example,
in weather data, temperature values might go missing due to their own high
temperature.  Or similarly, other measurements go missing due to high
temperatures.  Identifying this requires exploring the data, and the
analysis path to take is obvious, as missing values are often omitted from
plots and other statistical summaries.  To understand their structure and
possible mechanisms, the analyst must explore the data with visualizations,
summaries, and modelling, in an iterative fashion.

While there have been many software tools for exploring missing data, they
do not work together.  The graphics literature provides several solutions
for exploring missings visually by incorporating them into the plot in some
way.  For example, imputing values to be 10\% below the minimum to include
all observations into a scatter plot \citep{Cook2007}, or treating missing
values as an equal category of data \citep{Unwin1996}.  The ideas from the
graphics literature need to be translated into tidy data tools to integrate
missing data handling in a data analysis pipeline. We have implemented these tools for exploring missing data into the \proglang{R} package \pkg{naniar} \citep{pkg:naniar}, available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=naniar}.

This paper is organized in the following way.  Section~\ref{background}
provides the background to tidy data principles and tools
(Section~\ref{tidy-data-concepts}) and missing data representations
(Section~\ref{missing-data-rep-dep}).  Section~\ref{existing-software} summarizes
existing software for handling missing data.  The new work extending the
tidy tools to facilitate exploring, visualizing, and imputing missing data
is discussed in Section~\ref{extensions}.  Graphics (Section~\ref{graphics}) and
Numerical summaries (Section~\ref{num-sum}) of missing values are then discussed.
An application illustrating the use of the new methods is shown in
Section~\ref{case-study}.  Section~\ref{discussion} discusses strengths,
limitations, and future directions.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{tidy-data-concepts}{%
\subsection{Tidy data concepts and methods}\label{tidy-data-concepts}}

Features of tidy data were formally described in \citet{Wickham2014}, and
were discussed in terms of their importance for data science by
\citet{Donoho2017}, and tools for data analysis.  Tidy data is defined by
\citet{r4ds} as:
%
% \begin{quote}
\begin{enumerate}
% \def\labelenumi{\arabic{enumi}.}
% \tightlist
\item
  each variable must have its own column.
\item
  each observation must have its own row.
\item
  each value must have its own cell.
\end{enumerate}
% \end{quote}
%
Tidy data is easier to work with and analyze because the variables are in
the same format as they would be put into modelling software.  This helps
the analyst works swiftly and clearly, closing up opportunities for errors.
Tidy data principles are general, but are comprehensively implemented in the
\proglang{R} programming language, so this paper focuses on \proglang{R}
\citep{rcore}.

Tidy tools require the same tidy data input and output.  This consistency
means multiple tools can be composed together into a sequence, allowing for
rapid, elegant, and complex operations.  Contrasting tidy tools are messy
tools.  These have tidy input but messy output.  Messy tools slow down
analysis by shifting the focus from analysis to transforming output so it is
the right shape for the next step in the analysis.  This makes the work at
each step harder to predict, and more complex and difficult to maintain.
This disrupts workflow, and invites errors.  Tidy tools fall into three
broad categories: data manipulation, visualization, and modelling.

\hypertarget{tidy-data-manip}{%
\subsubsection{Data manipulation}\label{tidy-data-manip}}

Data manipulation is made input- and output-tidy with \proglang{R}~packages
\pkg{dplyr} and \pkg{tidyr} \citep{dplyr, tidyr}.  These provide the five
``verbs'' of data manipulation: data reshaping, sorting, filtering,
transforming, and aggregating.  Data reshaping goes from long to wide
formats; sorting arranges rows in a specific order; filtering removes rows
based on a condition; transforming, changes existing variables or adds new
ones; aggregating creates a single value from many values, say, for example,
in computing the minimum, maximum, and mean.

\hypertarget{tidy-vis}{%
\subsubsection{Visualizations}\label{tidy-vis}}

Visualization tools only have tidy data as their input, as the output is a
graphic.  The popular domain specific language \pkg{ggplot2} maps variables
in a dataset to features (referred to as aesthetics) of a graphic
\citep{ggplot2}.  For example, a scatterplot can be created by mapping two
variables to the $x$ and $y$~axes, and specifying a point geometry.

\hypertarget{tidy-model}{%
\subsubsection{Modelling}\label{tidy-model}}

Modelling tools work well with tidy data, as they have a clear mapping from
variables in the data to the formula for a model.  For example in
\proglang{R}, $y$ regressed on $x$ and $z$ is: \code{lm(y ~ x + z)}.  Modelling
tools are input tidy, but their output is always messy -- it is not in the
right format for subsequent steps in analysis.  For example, estimated
coefficients, predictions, and residuals from one model cannot be easily
combined with the output of another model.  Messy models have been partially
addressed with the \pkg{broom} package, which tidies up model outputs into a
tidy data format for data analysis, and the developing \pkg{recipes}
package, which helps make modelling input- and output-tidy \citep{recipes}.

\hypertarget{the-tidyverse}{%
\subsubsection{Tidy data and tidy tools}\label{the-tidyverse}}

Defining tidy data and tidy tools has resulted in a growing set of packages
known collectively as the \pkg{tidyverse} \citep{tidyverse}.  These are
constructed to share similar principles in their design and behavior, and
cover the breadth of an analysis - from importing, tidying, transforming,
visualizing, modelling, to communicating \citep{r4ds, tidyverse}.  This has
led to more tools for specific parts of analysis - from reading in data with
\pkg{readr}, \pkg{readxl}, and \pkg{haven}, to handling character strings
with \pkg{stringr}, dates with \pkg{lubridate}, and performing functional
programming with \pkg{purrr} \citep{readr, readxl, haven, stringr,
lubridate, purrr}.  It has also led to a burgeoning of new packages for
other fields following similar design principles, creating fluid workflows
for new domains.  For example, the \pkg{tidytext} \citep{tidytext} package
for text analysis, the \pkg{tsibble} \citep{wang2020tsibble} package for
time series data, and \pkg{tidycensus} \citep{tidycensus} for working with
US census and boundary data.

\hypertarget{tidy-formats-missing-data}{%
\subsubsection{Tidy formats for missing data}\label{tidy-formats-missing-data}}

Current tools for missing data are messy.  Missing data tools can be used to
perform imputations, missing data diagnostics, and data visualizations.
However, these tools suffer the same problems as modelling: they use tidy
input, but produce messy output -- their output is challenging to integrate
with other steps of data analysis.  The complex, often multivariate nature
of imputation methods also makes them difficult to represent.  Visualization
methods for missing data do not map data features to the aesthetics of a
graphic, as in \pkg{ggplot2}, limiting expressive exploration.

Taking existing methods from the missing data graphics literature, and
translating and expanding them into tidy data and tidy tools would create
more effective data visualizations.  Defining these concepts allows the
focus to be more general than just software, but rather, an extensible
framework for tidy tools to explore missing data.

\hypertarget{missing-data-rep-dep}{%
\subsection{Missing data representation and dependence}\label{missing-data-rep-dep}}

The convention for representing missingness is a \textbf{b}inary matrix,
\(B\), for data \(y\) with \(i\) rows and \(j\) columns:
%
\[
b_{ij} =\begin{cases}
1, & \text{if } y_{ij} \text{ is missing} \\
0, & \text{if } y_{ij} \text{ is observed}
\end{cases}
\]
%
There are many ways each value can be missing, we adopt the notation used in
\citep{vanBuuren2018}.  The information in \(B\) can be used to arrive at
three categories of missing values: missing completely at random (MCAR),
missing at random (MAR), and missing not at random (MNAR).  The distribution
of missing values in \(b_{ij}\) can depend on the entire dataset,
represented as \(Y = (Y_\text{obs}, Y_\text{miss})\).  This relationship can be
defined by the \emph{missing data model} \(\Prob(b_{ij} \mid Y_\text{obs}, Y_\text{miss},
\psi)\), the probability of missingness is conditional on data observed,
data missing, and some probability parameter of missingness, \(\psi\).  This
helps to precisely define categories of missing values.

MCAR is where values being missing have no association with
observed or unobserved data, that is, \(\Prob(B = 1 \mid Y_\text{obs}, Y_\text{miss}) = \Prob(B
= 1 \mid \psi)\).  Essentially, the probability of an observation being missing
is unrelated to anything else, only the parameter \(\psi\), the overall
probability of missingness.  Although a convenient scenario, it is not
actually possible to confirm, or clearly distinguish from MAR, as it relies
on statements on data unobserved.  In MAR, missingness only depends
on data observed, not data missing, that is, \(\Prob(B = 1 \mid Y_\text{obs}, Y_\text{miss},
\psi) = \Prob(B = 1 \mid Y_\text{obs}, \psi)\).  Some structure or dependence between
missing and observed values is allowed, provided it can be explained by data
observed, and some overall probability of missingness.  In MNAR,
missingness is related to values observed, and unobserved: \(\Prob(B = 1 \mid
Y_\text{obs}, Y_\text{miss}, \psi)\).  This assumes conditioning on all observations:
data goes missing due to some phenomena unobserved, including the structure
of the missing data itself.  This presents a challenge in analysis, as it is
difficult to verify, and implies bias in analysis due to the unobserved
phenomena.  Visualizations can help assess whether data may be MCAR, MAR or
MNAR.

\hypertarget{existing-software}{%
\section{Existing Software}\label{existing-software}}

Methods for exploring, understanding, and imputing missing data are more
accessible now than they have ever been.  Values can be imputed with one
value (single imputation), or multiple values (multiple imputation),
creating \(m\) datasets.  This section discusses existing software for
single and multiple imputation, and missing data exploration.

\hypertarget{imputation}{%
\subsection{Imputation}\label{imputation}}

\pkg{VIM} \citep{VIM} implements well-used imputation methods: $K$ nearest
neighbors, regression, hot-deck, and iterative robust model-based
imputation.  These diverse approaches allow for imputing with
semi-continuous, continuous, count, and categorical data.  \pkg{VIM} identifies
imputed cases by adding an indicator variable with a suffix \code{_imp}.
So, \code{Var1} has a sibling column, \code{Var1_imp}, with values
\code{TRUE} or \code{FALSE} indicating imputation.  \pkg{VIM} also has a variety of
visualization methods, discussed in Section~\ref{exploration}.
\pkg{simputation} provides an interface to imputation methods from
\pkg{VIM}, in addition providing hotdeck imputation, and the
Expectation-maximization (EM) algorithm
\citep{simputation, Dempster1977}.  \pkg{simputation} provides a consistent
formula interface for all imputation methods, and always returns a dataframe
with the updated imputations.  \pkg{Hmisc} \citep{Hmisc}, provides
predictive mean matching, \pkg{imputeTS} \citep{imputeTS}, provides time
series imputation methods, and \pkg{missMDA} \citep{missMDA}, imputes data
using principal component analysis.

Multiple imputation is often regarded as best practice for imputing values
\citep{Schafer2002}, as long as appropriate caution is taken
\citep{Sterne2009}.  Popular and robust methods for multiple imputation
include the \pkg{mice}, \pkg{Amelia}, and \pkg{mi} packages \citep{mice,
amelia, mi}.  \pkg{mice} implements the method of chained equations, using a
variable-wise algorithm to calculate the posterior distribution of
parameters to generate imputed values.  The workflow in \pkg{mice} revolves
around imputing data, returning completed data, and fitting a model and
pooling the results.

\pkg{Amelia} \citep{amelia} assumes data are multivariate normal, and
samples from the posterior, and allows for incorporation of information on
the values in a prior.  It uses the computationally efficient (and
parallelizable) Expectation-Maximization Bootstrap (EMB) algorithm
\citep{Honaker2010}.  \pkg{norm} \citep{norm}, provides multiple imputation
using EM for multivariate normal data, drawing from methods in the
\proglang{NORM} software \citep{schafer-norm}.  \pkg{norm} does not provide
a framework for tracking missing values, instead providing tools for making
inference from multiple imputation.

\pkg{mi} \citep{mi} also uses Bayesian models for imputation, providing
better handling of semi-continuous values, and data with structural or
perfect correlation.  A collection of analysis models are also provided in
\pkg{mi}, to work with data it has imputed.  These include linear models,
generalized linear models, and their corresponding Bayesian components.
This approach promotes fluid workflow, with a similar penalty to tidying up
model output, which is still messy.

\hypertarget{imputation-summary}{%
\subsubsection{Summary}\label{imputation-summary}}

Each imputation method provides practical methods for different use cases,
but most have different output structures, and do not have consistent
interfaces in their implementation.  This makes them inherently messy and
challenging to integrate into an analysis pipeline.  For example, combining
different imputation methods from different pieces of software is not
currently straightforward.  \pkg{simputation} resolves some of these
complications with a simple approach of a unified syntax for all imputation,
and always returns a dataframe of imputed values.  This reduces the friction
of working with other tools, but comes at the cost of identifying imputed
values.  An ideal approach would use consistent, simple data structures that
work with other analysis tools, and help track missing values.  This would
make imputation outputs tidy, streamlining subsequent analysis.

\hypertarget{exploration}{%
\subsection{Exploration}\label{exploration}}

Missing data packages are focused primarily on making inferences, and
exploring imputed values, not on exploring relationships in missing values,
and identifying possible patterns.  Texts covering the exploration phase of
missing data have the same problem as with modelling: the input is tidy, but
the output does not work with other tools \citep{vanBuuren2018}; this is
inefficient.  Methods for exploring missing values are primarily covered in
literature on interactive graphics \citep{Swayne1998, Unwin1996, Cook2007},
and are picked up again in a discussion of a graphical user interface
\citep{Cheng2015}.

The missingness matrix \(B\) can be used to assess missing data dependence.
It has been used in interactive graphics, dubbed a ``shadow matrix'', to
link missing and imputed values to the data, facilitating their display
\citep{Swayne1998}, focusing heavily on multivariate numeric data.  This is
an idea upon which this new work builds.

The \proglang{MANET} (missings are now equally treated) software
\citep{Unwin1996} focused on multivariate categorical data, with
missingness explicitly added as a category.  \proglang{MANET} also provided
univariate visualizations of missing data using linked brushing between a
reference plot of the missingness for each variable, and a plot of the data
as a histogram or barplot.  The \proglang{MANET} software is no longer
maintained and cannot be installed.  The approach of \citep{Swayne1998} in
the software \proglang{XGobi}, further developed in \proglang{ggobi}
\citep{Cook2007}, focused on multivariate quantitative data.  Missingness
is incorporated into plots in \proglang{ggobi} by setting them to be 10\%
below the minimum value.

\pkg{MissingDataGUI} \citep{pkg:MissingDataGUI}
provides a Graphical User Interface (GUI) for exploring
missing data structure, both numerically and visually.  Using a GUI to
explore missing data facilitates rapid insight into missingness structures.
However, this comes as a trade off, as insights are not captured or recorded
with a GUI, making it challenging to incorporate into reproducible analyses.
This distracts and breaks analysis workflow, inviting mistakes.

\pkg{VIM} (visualizing and imputing missing data) provides visualization
methods to identify and explore observed, imputed, and missing values.
These include spinograms, spinoplots, missingness matrices, plotting
missingness in the margins of other plots, and other summaries.  However,
these visualizations do not map variables to graphical aesthetics, creating
friction when moving through analysis workflows, making them difficult to
extend to new circumstances.  Additionally, data used to create the
visualizations cannot be accessed, posing a barrier to further exploration.

The \pkg{inspectdf} package provides tools for summarizing, comparing, and
visualizing dataframes \citep{inspectdf}.  In particular the
\code{inspect\_na()} function provides a summary of missing values for
variables.  \code{inspect\_cat()} summarizes the levels of categories in
each column in a dataset.  The \code{show\_plot()} function shows a higher
lever overview of the contents of a dataframe and its categories, as well as
missing values.  All of these functions allow for comparing missing values
across two dataframes.  For example, comparing the missing values in two
dataframes with the same column names.

\pkg{ggplot2} removes missing values with a warning
(Figure~\ref{fig:warning}), and only incorporates missingness into
visualizations when mapping a discrete variable with missings to a graph
aesthetic.  This has some limitations, shown in Figure~\ref{fig:gg-box-na},
a boxplot visualization of school grade and test scores.  If there are
missings in a continuous variable like test score, \pkg{ggplot2} omits the
missings and prints a warning message.  However, if a discrete variable like
school year has missing values, an NA category is created for school year,
where scores are placed.

\begin{figure}[t!]
\centering
\includegraphics[width=1\linewidth]{gg-box-na-1}
%
\caption[\pkg{ggplot2} provides different visualizations depending on what
type of data has missing values for data of student test scores in school
year]{\pkg{ggplot2} provides different visualizations depending on what type
of data has missing values for data of student test scores in school year.
(A) Data is complete; (B) Missings are only in year -- an additional ``NA''
boxplot is created; (C) Missings only in scores, no additional missingness
information is shown; (D) Missings in both scores and year, additional
missing information is shown.  The missingness category is only shown when
there are missings in categorical variables such as year (plots (B) and
(D)).  In (C), no missingness information is given on the graphic, despite
there being missings in score, and a warning message is displayed about the
number of missing values omitted.}\label{fig:gg-box-na}
%
\end{figure}

\hypertarget{approaches-in-other-languages}{%
\subsection{Approaches in other languages}\label{approaches-in-other-languages}}

Missing data can be explored and visualized in closed source software such
as \proglang{STATA}, \proglang{SAS}, and \proglang{SPSS}.  These provide
basic table frequencies of missing data, and basic visualizations on
missingness patterns.  However, none of these software provide functions to
specifically visualize missing values in bivariate or multivariate settings,
and it is only possible to visualize missing values after performing data
analysis, counting and summarizing missingness.  Imputation methods from
mean to regression, EM, and multiple imputation are all available, with out
of the box diagnostics provided to assist comparing imputations.
\proglang{Python} also provides exploratory data analysis for missing data
within \pkg{missingno} \citep{Bilogur2018}, and the \pkg{scikit-learn}
package provides imputation methods \citep{scikit-learn}.

\hypertarget{extensions}{%
\section{Tidy framework for missings}\label{extensions}}

Applying \pkg{tidyverse} principles has the potential to clarify missing
data exploration, visualization, and imputation.  This section discusses how
these principles are applied for data structures (Section~\ref{data-structure}),
common operations (verbs) (Section~\ref{verbs}), graphics (Section~\ref{graphics}), and data
summaries (Section~\ref{num-sum}).  Care has been taken to make the names and design
these features intuitive for their purpose, and is discussed throughout.

\hypertarget{data-structure}{%
\subsection{Data structure}\label{data-structure}}

A data structure facilitating exploration of missing data needs to be simple
to reason with and to transport with a data analysis, otherwise it will not
be used.  A useful template common in missing data literature, is the \(B\)
matrix, where 0 and 1 indicate not missing and missing, respectively.  This
matrix was used to explore missing values in the interactive graphics
library \proglang{XGobi}, called a ``missing value shadow'', or ``shadow
matrix'', defined as a copy of the original data with indicator values of
missingness.  The shadow matrix could be interactively linked to the data.
However, there are some limitations to the shadow matrix.  Namely, the
values 0 and 1 can be confusing representations of missing values, since it
is not clear if 0 indicates an absence of observation, or the presence of a
missing value.

We propose a new form for tidy representation of missing data based on these
ideas from past research.  Four features are added to the shadow matrix to
facilitate analysis of missing data, illustrated in
Figure~\ref{fig:nabularfig}.  These are briefly described below, and
discussed in further detail in Section~\ref{nabular-data}.
%
\begin{enumerate}
% \def\labelenumi{\arabic{enumi}.}
%
\item
  \textbf{Missing value labels}:  Simple labels for missing and not missing
to clearly identify missing values for analysis and plotting.  We propose
``NA'' and ``!NA'' (Figure~\ref{fig:nabularfig}).  This improves the 0 and
1 values in \(B\), which do not clearly identify missingness.  These values
follow a principle: ``clarity of labelling'' -- the matrix's meaning
is transparent, and anybody looking at these values could understand what
they mean, which is not the case of binary values.  Equally, these values
could instead be ``missing'' or ``present''.
%
\item
  \textbf{Special missing values}: Building on \textbf{missing value
labels}, the values in the shadow matrix can be ``special'' missing values,
indicated by ``NA\_\textless suffix\textgreater{}''.  For example:
``NA\_instrument'' uses a short label, ``instrument'', indicating instrument
error resulting in missing values.  These could be also used to indicate
imputations (Figure~\ref{fig:nabularfig}).
%
\item
  \textbf{Coordinated names}: Variable names in the shadow matrix gain a
consistent short suffix, ``\_NA'', keeping names coordinated throughout
analysis (Figure~\ref{fig:nabularfig}).  It makes a clear distinction with
``var\_NA'' being a random variable of the missingness of a variable,
``var''.  This suffix is short and easy to remember during data analysis,
and shifts the focus from the value of a variable, to its missingness state.
Although this might break naming conventions for lower case ``snake\_case''
variables \citep[e.g.,][]{Wickham2014}, this distinguishes these columns from
other columns.
%
\item
  \textbf{Connectedness}: Binding the shadow matrix column-wise to the
original data as a factor creates a single, connected, nabular data format,
in sync with the data, at the cost of a larger file.  It is useful for
visualization, summaries, and tracking imputed values, discussed in more
detail in Section~\ref{nabular-data}.
\end{enumerate}

\hypertarget{nabular-data}{%
\subsection{Nabular data}\label{nabular-data}}

\emph{Nabular} data binds the shadow matrix column-wise to the original
data.  It is a portmanteau of ``NA'' and ``tabular''.  \emph{Nabular} data
explicitly links missing values to data, keeping corresponding observations
together and removing the possibility of mismatching records
(Figure~\ref{fig:nabularfig}).  \emph{Nabular} data facilitates
visualization and summaries by allowing the user to reference the
missingness of a variable in a coordinated way: the missingness of ``var'',
as ``var\_NA'' during analysis.  \emph{Nabular} data is a snapshot of the
missingness of the data.  This means when \emph{nabular} data are imputed,
those imputed values can easily be identified in analysis (``var'' vs
``var\_NA'').  Imputing values on \emph{nabular} data automatically tracks
these imputations.  Together these features make it possible to do tasks
that are not possible by simply adding a binary missing matrix, \(B\) during
data analysis.

\begin{figure}[t!]
\centering
\includegraphics[width=1\linewidth]{diagram}
\caption[The process of creating nabular data]{The process of creating
nabular data.  Data transformed to shadow matrix, where values are either
not missing or missing: ``!NA'' or ``NA''.  The shadow matrix can be converted
to long form to create missingness summary plots.  Nabular data is created
by binding the columns of the data and shadow matrix.  Special missing
values (such as ``$-99$'') are identified as special missings, and values imputed
and tracked.  Nabular data can be used to identify imputations and explore
data values alongside missings, providing a useful format for missing data
exploration and analysis.}\label{fig:nabularfig}
\end{figure}

\emph{Nabular} data is similar to classical data formats with quality or
flag variables associated with each measured variable, e.g.,~Scripps CO2
data \citep{Keeling2005-scripps}, GHCN data \citep{Durre2008-ghcn}.  Using
additional columns to represent missingness information follows best
practices for data organization, described in \citep{Ellis2018} and
\citep{Broman2018}: (1) keep one thing in a cell and (2) describe additional
features of variables in a second column.  Here they suggest to indicate
censored data with an extra variable called ``VariableNameCensored'', which
would be TRUE if censored, otherwise FALSE.  This information can now be
represented in the shadow columns as special missing values.  Encoding
special missing values is achieved by defining logical conditions and
suffixes.  This is implemented with the \code{recode_shadow()} function in
\pkg{naniar} (discussed in Section~\ref{verbs}).

Special missing values are not a new idea, and have been implemented in
other statistical programming languages, \proglang{SPSS}, \proglang{SAS},
and \proglang{STATA}.  These typically represent missing values as a
full-stop, ``.'', and record special missing values as ``.a'' -- ``.z''.
These special values from these languages break the tidy principle of one
column having one type of value, as they record both the value, and the
multivariate missingness state.

\hypertarget{verbs}{%
\subsection{Missing data operations}\label{verbs}}

Common missing data operations can be considered verbs, in the tidyverse
sense.  For missing data, these include: \textbf{scan}, \textbf{replace},
\textbf{add}, \textbf{shadow}, \textbf{impute}, \textbf{track}, and
\textbf{flag}.  Data can be \textbf{scanned} to find possible missings not
coded as NA.  These values can then be \textbf{replaced} with
NA.  To facilitate exploration, summaries of missingness can be
\textbf{added} as a summary column to the original data.  The data can be
augmented with the \textbf{shadow matrix} values, helping explore missing
data, as well as facilitating the process of \textbf{imputing}, and
\textbf{tracking}.  Finally, unusual or specially coded missing values can
be \textbf{flagged}.

\hypertarget{verbs-search}{%
\subsubsection{``scan'': Searching for common missing value labels}\label{verbs-search}}

This operation is used to search the data for specific conventional
representations of missings, such as, ``N/A'', ``MISSING'', ``$-99$''.  This
is implemented in the function \code{miss_scan_count()}, which returns a
table of occurrences of that value for each variable.  A list of common NA
values for numbers and characters, can be provided to help check for typical
representations of missings.  These are implemented in \pkg{naniar} as
objects `\code{common_na_numbers}' and `\code{common_na_strings}'.

\hypertarget{verbs-replace-with}{%
\subsubsection{``replace'': Replacing values with missing values}\label{verbs-replace-with}}

Once possible missing values have been identified, these values can be
replaced.  For example, a dataset could have the values ``$-99$'' meaning a
missing value.  \pkg{naniar} implements replacement with the function,
\code{replace_with_na()}.  Values ``$-99$'' could be replaced in the ``x''
column with: \code{replace_with_na(dat_ms, replace = list(x = -99))}.  For
operating on multiple variables, there are scoped variants for
\code{replace_with_na()}: \code{_all}, \code{_if}, and \code{_at}.  This
means \code{replace_with_na_all()} operates on \textbf{all} columns,
\code{replace_with_na_at()} operates \textbf{at} specific columns, and
\code{replace_with_na_if()} makes a conditional change on columns
\textbf{if} they meet some condition (such as \code{is.numeric()} or
\code{is.character()}).

\hypertarget{verbs-add-cols}{%
\subsubsection{``add'': Adding missingness summary variables}\label{verbs-add-cols}}

Understanding the missingness structure can be improved by adding summary
information alongside the data.  For example, in \citep{Tierney2015}, the
proportion of missings in a row is used as the outcome in a model to
identify variables important in predicting missingness structures.
\pkg{naniar} implements a series of functions to add these missingness
summaries to the data, starting with \code{add_}.  These are inspired by the
\code{add_count()} function in \pkg{dplyr}, which adds count information for
specified groups or conditions.  \pkg{naniar} provides operations to add the
number or proportion of missingness, the missingness cluster, or the
presence of any missings, with: \code{add_n_miss()}, \code{add_prop_miss()},
\code{add_miss_cluster()}, and \code{add_any_miss()}, respectively
(Table~\ref{tab:add-missing-info}).  There are also functions for adding
information about shadow values, and readable labels for any missing values
with \code{add_label_shadow()} and \code{add_label_missings()}.

\begin{table}[t!]
\centering
\begin{tabular}[t]{ll}
\hline
Function & Adds column which\\
\hline
\code{add\_n\_miss(data)} & contains the number missing values in a row\\
\code{add\_any\_miss(data)} & contains whether there are any missing values in a row\\
\code{add\_prop\_miss(data)} & contains the proportion of missing values in a row\\
\code{add\_miss\_cluster(data)} & contains the missing value cluster\\
\hline
\end{tabular}
\caption{\label{tab:add-missing-info}Overview of the \code{add} functions in
\pkg{naniar}}.
\end{table}

\hypertarget{verbs-nabular}{%
\subsubsection{``shadow'': Creating nabular data}\label{verbs-nabular}}

\emph{Nabular} data has the shadow matrix column-bound to existing data.
This facilitates visualization and summaries, and allows for imputed values
to be tracked.  \emph{Nabular} data can be created with \code{nabular()}:

\begin{CodeChunk}
\begin{CodeInput}
R> nabular(dat_ms)
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 5 x 6
      x y         z x_NA  y_NA  z_NA
  <dbl> <chr> <dbl> <fct> <fct> <fct>
1     1 A      -100 !NA   !NA   !NA
2     3 N/A     -99 !NA   !NA   !NA
3    NA <NA>    -98 NA    NA    !NA
4   -99 E      -101 !NA   !NA   !NA
5   -98 F        -1 !NA   !NA   !NA
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{verbs-recode}{%
\subsubsection{``flag'': Describing different types of missing values}\label{verbs-recode}}

Unusual or spurious data values are often identified and ``flagged''.  For
example, there might be special codes to mark an individual dropping out of
a study, known instrument failure in weather instruments, or for values
censored in analysis.  \pkg{naniar} provides tools to encode these special
types of missingness in the shadow matrix of \emph{nabular} data, using
\code{recode_shadow()}.  This requires specifying the variable to contain
the flagged value, the condition for flagging, and a suffix.  This is then
recoded as a new factor level in the shadow matrix, so every column is aware
of all possible new values of missingness.  For example, ``$-99$'' could be
recoded to indicate a broken machine sensor for the variable ``x'' with the
following:
%
\begin{CodeChunk}
\begin{CodeInput}
R> nabular(dat_ms) %>%
+    recode_shadow(x = .where(x == -99 ~ "broken_sensor"))
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 5 x 6
      x y         z x_NA             y_NA  z_NA
* <dbl> <chr> <dbl> <fct>            <fct> <fct>
1     1 A      -100 !NA              !NA   !NA
2     3 N/A     -99 !NA              !NA   !NA
3    NA <NA>    -98 NA               NA    !NA
4   -99 E      -101 NA_broken_sensor !NA   !NA
5   -98 F        -1 !NA              !NA   !NA
\end{CodeOutput}
\end{CodeChunk}
%
\hypertarget{verbs-impute}{%
\subsubsection{``impute'': Imputing values}\label{verbs-impute}}

\pkg{naniar} does not reinvent the wheel by creating the need for customized
imputation procedures.  Instead, the \code{nabular} format means existing
methods using imputation replacement in place now automatically track
missing values.  \pkg{naniar} provides a few imputation methods to
facilitate exploration and visualization: \code{impute_below()},
\code{impute_mean()}, and \code{impute_median()}.  While useful to explore
structure in missingness, they are not recommended for use in analysis.
\code{impute_below()} imputes values below the minimum value, with some
controllable jitter (random noise) to reduce overplotting.

Similar to \pkg{simputation}, each \code{impute_} function returns the data
with values imputed.  However, \pkg{naniar} does not use a formula syntax,
instead each function has ``scoped variants'' \code{_all}, \code{_at} and
\code{_if}.  \code{impute_} functions with no scoped variant,
(\code{impute_mean()}), will work on a single vector, but not a data.frame.
One challenge with this approach is imputed value locations are not tracked.
This issue is resolved with \emph{nabular} data covered in the following
section.

\hypertarget{verbs-track}{%
\subsubsection{``track'': Shadow and impute missing values}\label{verbs-track}}

To evaluate imputations they need to be tracked.  This is achieved by first
using \emph{nabular} data, then imputing, and imputed values can then be
referred to by their shadow variable, ``\_NA''
(Figure~\ref{fig:track-impute-example}).  The code below shows the track
pattern, first using \code{nabular()}, then imputing with
\code{impute_lm()}.  \code{label_shadow()} then adds a label to facilitate
identifying missings:
%
\begin{CodeChunk}
\begin{CodeInput}
R> aq_imputed <- nabular(airquality) %>%
+   as.data.frame() %>%
+   simputation::impute_lm(Ozone ~ Temp + Wind) %>%
+   simputation::impute_lm(Solar.R ~ Temp + Wind) %>%
+   add_label_shadow()
R> head(aq_imputed)
\end{CodeInput}
\begin{CodeOutput}
      Ozone  Solar.R Wind Temp Month Day Ozone_NA Solar.R_NA Wind_NA Temp_NA
1  41.00000 190.0000  7.4   67     5   1      !NA        !NA     !NA     !NA
2  36.00000 118.0000  8.0   72     5   2      !NA        !NA     !NA     !NA
3  12.00000 149.0000 12.6   74     5   3      !NA        !NA     !NA     !NA
4  18.00000 313.0000 11.5   62     5   4      !NA        !NA     !NA     !NA
5 -11.67673 127.4317 14.3   56     5   5       NA         NA     !NA     !NA
6  28.00000 159.5042 14.9   66     5   6      !NA         NA     !NA     !NA
  Month_NA Day_NA any_missing
1      !NA    !NA Not Missing
2      !NA    !NA Not Missing
3      !NA    !NA Not Missing
4      !NA    !NA Not Missing
5      !NA    !NA     Missing
6      !NA    !NA     Missing
\end{CodeOutput}
\end{CodeChunk}
%
Multiple missing or imputed values can be mapped to a graphical element in
\pkg{ggplot2} by setting the \code{color} or \code{fill} aesthetic in ggplot to
\code{any_missing}, a result of the \code{add_label_shadow()} function.
(Figure~\ref{fig:track-impute-example}).  Imputed values can also be
compared to complete case data, grouping by \code{any_missing}, and then
summarizing, similar to other \pkg{dplyr} summary workflows shown below in
Table~\ref{tab:impute-summary-out}, showing similarities and differences in
imputation methods.

\begin{figure}[t!]
\centering
\includegraphics[width=1\linewidth]{track-impute-example-1}
\caption[Scatterplot (A) and density plots (B and C) of ozone and solar
radiation from the airquality dataset containing imputed values from a
linear model]{Scatterplot (A) and density plots (B and C) of ozone and solar
radiation from the airquality dataset containing imputed values from a
linear model.  Imputed values are colored green, and data values orange.
Imputed values are similar, but slightly trended to the
mean.}\label{fig:track-impute-example}
\end{figure}

\begin{CodeChunk}
\begin{CodeInput}
R> aq_imputed %>%
+   group_by(any_missing) %>%
+   summarise_at(.vars = vars(Ozone), .funs = lst(min, mean, median, max))
\end{CodeInput}
\end{CodeChunk}

\begin{table}[t!]
\centering
\begin{tabular}[t]{lrrrr}
\hline
\code{any\_missing} & \code{min} & \code{mean} & \code{median} & \code{max}\\
\hline
Missing & $-16.86418$ & 41.22494 & 45.4734 & 78\\
Not Missing & 1.00000 & 42.09910 & 31.0000 & 168\\
\hline
\end{tabular}
\caption{\label{tab:impute-summary-out}Output of \pkg{dplyr} summary
statistics of imputed vs non imputed values for the variable ``Ozone''.  The
``any\_missing'' column denotes imputed values (``Missing'', since they were
previously missing), and non-imputed values (``Not Missing'').  The mean and
median values are similar, but the minimum and maximum values are very
different.}
\end{table}

\hypertarget{graphics}{%
\section{Graphics}\label{graphics}}

Missing values are often ignored when plotting data - which is why the data
visualization software, \proglang{MANET}, was named and is an acronym
corresponding to ``missings are now equally treated'' \citep{Unwin1996}.
However, plots can help to identify the type of missing value patterns, even
those of MCAR, MAR or MNAR.  Here we summarize how to systematically explore
missing patterns visually, and define useful plots to make, relative to the
\emph{nabular} data structure.

\hypertarget{overviews}{%
\subsection{Overviews}\label{overviews}}

The first step is to get an overview of the extent of the missingness.
Overview visualizations for variables and cases are provided with
\code{gg_miss_var()} and \code{gg_miss_case()}
(Figure~\ref{fig:gg-miss-case-var}A), drawing attention to the amount of
missings, and ordering by missingness.  The ``airquality'' dataset (from
base \proglang{R}), is shown, and contains daily air quality measurements in
New York, from May to September, 1973.  We learn from
Figure~\ref{fig:gg-miss-case-var}A--B, that two variables contain missings,
approximately one third of observations have one missing value, and a tiny
number of observations are missing across two variables.  These overview
plots are created from the shadow matrix in long form
(Figure~\ref{fig:nabularfig}).  Numerical statistics can also be reported
(Section~\ref{num-sum}).

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{gg-miss-case-var-1}
\caption[Graphical summaries of missingness in the airquality
data]{Graphical summaries of missingness in the airquality data.  Missings
in variables (A) and cases (B), and for an overview of all missingness as a
heatmap in (C), and with clustering applied (D).  There are missing values
in ``Ozone'' and ``Solar.R'', with Ozone having more missings.  Not many cases have
two missings.  Most missingness is from cases with one missing value.  The
default output (C) and ordered by clustering on rows and columns (D).  These
overviews are made possible using the shadow matrix in long form.  There are
only missings in ozone and solar radiation, and there appears to be some
structure to their missingness.}\label{fig:gg-miss-case-var}
\end{figure}

The shadow matrix can be put into long form, allowing both the variables and
cases to be displayed using a heatmap style visualization, with
\code{vis_miss()} from \pkg{visdat} \citep{visdat}
(Figure~\ref{fig:nabularfig}).  This also provides numerical summaries of
missingness in the legend, and for each column
(Figure~\ref{fig:gg-miss-case-var}C--D).  Clustering can be applied to the
rows, and columns arranged in order of missingness
(Figure~\ref{fig:gg-miss-case-var}D).  Similar visualizations are available
in other packages such as \pkg{VIM}, \pkg{mi}, \pkg{Amelia}, and
\pkg{MissingDataGUI}.  A key improvement is \code{vis_miss()} orients the
visualization analogous to a regular data structure: variables form columns
and are named at the top, and each row is an observation.  Using
\pkg{ggplot2}, as the foundation, makes the plot easily customizable.

\begin{figure}
\centering
\includegraphics{airquality-upset-1}
\caption[The pattern of missingness in the airquality dataset shown in an
upset plot]{The pattern of missingness in the airquality dataset shown in an
upset plot.  Only ``Ozone'' and ``Solar.R'' have missing values, and Ozone has the
most missing values.  There are 2~cases where both ``Solar.R'' and Ozone have
missing values.}\label{fig:airquality-upset}
\end{figure}

The number of times observations are missing together can be visualized
using an ``upset plot'' \citep{Conway2017}.  An alternative to a Venn
diagram, an upset plot shows the size and features of overlapping sets, and
scales well with more variables.  An upset plot can be constructed from the
shadow matrix, as shown in Figure~\ref{fig:airquality-upset} which displays
the overlapping counts of missings in the airquality data.  The bottom right
shows the combinations of missingness, the top panel shows the size of these
combinations, and the bottom left shows missingness in each variable.  This
provides similar information to Figure~\ref{fig:gg-miss-case-var}A--D, but
more clearly illustrating overlapping missingness, where 2~cases are missing
together in variables ``Solar.R'' and ``Ozone''.

\hypertarget{univariate}{%
\subsection{Univariate}\label{univariate}}

Missing values are typically not shown for univariate visualizations such as
histograms or densities.  Two ways to use \emph{nabular} data to present
univariate data with missings are discussed.  The first imputes values below
the range to facilitate visualizations.  The second displays two plots of
the same variable according to the missingness of a chosen variable.

\subsubsection{Imputing values below the range}

To visualize the amount of
missings in each variable, the data is transformed into \emph{nabular} form,
then values are imputed below the range of data using
\code{impute_below_all()}.  Figure~\ref{fig:impute-shift-histogram}A shows a
histogram of Ozone values on the right in green, and the histogram of
missing ozone values on the left, in orange.  The missings in Ozone are
imputed and ``Ozone\_NA'' is mapped to the fill aesthetic.  \emph{Nabular}
data facilitates adding counts of missingness to a histogram, allowing
examination of a variable's distribution of values, and also the magnitude
of missings.

\subsubsection{Univariate split by missingness}

The distribution of a variable
can be shown according to the missingness of another variable.  The shadow
matrix part of \emph{nabular} is used to handle the faceting, and color
mapping.  Figure~\ref{fig:impute-shift-histogram} shows the values of
temperature when ozone is present, and missing, using a faceted histogram
(B), and an overlaid density (C).  This shows how values of temperature are
affected by the missingness of ozone, and reveals a cluster of low
temperature observations with missing ozone values.  This type of plot can
facilitate exploring missing data distributions.  For example, we would
expect if data were MCAR, for values to be roughly uniformly missing
throughout the histogram or density.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\linewidth]{impute-shift-histogram-1}
\includegraphics[width=0.75\linewidth]{impute-shift-histogram-2}
\includegraphics[width=0.75\linewidth]{impute-shift-histogram-3}
\caption[Univariate summaries of missingness]{Univariate summaries of
missingness.  (A) A histogram using nabular data to show the values and
missings in ozone.  Values are imputed below the range to show the number of
missings in ozone and colored according to missingness of ozone
(``Ozone\_NA'').  There are about 35 missings in Ozone.  Panel~C shows
temperature according to missingness in ozone from in the airquality
dataset.  A histogram of temperature faceted by the missingness of ozone
(B), or a density of temperature colored by missingness in ozone (C).  These
show a cluster of low temperature observations with missing ozone values,
but temperature is otherwise similar.}\label{fig:impute-shift-histogram}
\end{figure}

\hypertarget{bivariate}{%
\subsection{Bivariate}\label{bivariate}}

To visualize missing values in two dimensions the missing values can be
placed in plot margins, by imputing values below the range of the data.
Using \emph{nabular} data identifies imputed values, and color makes
missingness pre-attentive \citep{treisman1985}.  The steps of imputing and
coloring are combined into \code{geom_miss_point()}.
Figure~\ref{fig:geom-miss} shows a mostly uniform spread of missing values
for ``Solar.R'' and ``Ozone''.  As \code{geom_miss_point()} is a defined
\pkg{ggplot2} geometry, it works with features such as faceting and mapping
other variables to graphical aesthetics.

\begin{figure}[t!]
\centering
\includegraphics{geom-miss-1}
\caption[Scatterplots with missings displayed at 10\% below for the
airquality dataset]{Scatterplots with missings displayed at 10\% below
for the airquality dataset.  Scatterplots of ozone and solar radiation (A),
and ozone and temperature (B).  There are missings in ozone and solar
radiation, but not temperature.}\label{fig:geom-miss}
\end{figure}

\hypertarget{multivariate}{%
\subsection{Multivariate}\label{multivariate}}

Parallel coordinate plots can help to visualize missingness beyond two
dimensions.  They transform variables to the same scale, ranging between 0
and 1.  The \code{oceanbuoys} dataset from \pkg{naniar} is used for this
visualization, containing measurements of sea and air temperature, humidity,
and east west and north south wind speeds.  Data was collected in 1993 and
1997, to understand and predict El Ni{\~n}o and El Ni{\~n}a.
Figure~\ref{fig:parallel-cord-plot} is a parallel coordinate plot of
\code{oceanbuoys}, with missing values imputed to be 10\% below the range,
and values colored according to whether humidity was missing
(``humidity\_NA'').  Figure~\ref{fig:parallel-cord-plot} shows humidity is
missing at low air and sea temperatures, and humidity is missing in one
year, and one location.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{parallel-cord-plot-1}
\caption[Parallel coordinate plot shows missing values imputed 10\% below
range for the oceanbuoys dataset]{Parallel coordinate plot shows missing
values imputed 10\% below range for the oceanbuoys dataset.  Values are
colored by missingness of humidity.  Humidity is missing for low air and sea
temperatures, and is missing for one year and one
location.}\label{fig:parallel-cord-plot}
\end{figure}

\hypertarget{num-sum}{%
\section{Numerical summaries}\label{num-sum}}

This section describes approaches to summarizing missingness, and an
implementation in \pkg{naniar}.  Numerical summaries should be easy to
remember with consistent names and output, returning either a single
number (Section~\ref{single-num-sum}), or a dataframe (Section~\ref{sum-tab-missings}),
so they integrate well with plotting and modelling tools.  How these work with other
tools in an analysis pipeline is shown in (Section~\ref{num-sum-w-group}).

\hypertarget{single-num-sum}{%
\subsection{Single number summaries}\label{single-num-sum}}

The overall number, proportion, or percent of missing values in a dataset
should be simple to calculate.  \pkg{naniar} provides the functions
\code{n_miss()}, \code{prop_miss()} and \code{pct_miss()}, as well as their
complements.  Summaries for variables and cases are made by appending
\code{_case} or \code{_var} to these summaries.  An overview is shown in
Table~\ref{tab:n-prop-pct-miss-complete}.

\begin{table}[t!]
\centering
\begin{tabular}[t]{lrlr}
\hline
Missing function & Missing value & Complete function & Complete value\\
\hline
\code{n\_miss} & 44.00 & \code{n\_complete} & 874.00\\
\code{prop\_miss} & 0.05 & \code{prop\_complete} & 0.95\\
\code{pct\_miss} & 4.79 & \code{pct\_complete} & 95.21\\
\code{pct\_miss\_case} & 27.45 & \code{prop\_complete\_case} & 72.55\\
\code{pct\_miss\_var} & 33.33 & \code{pct\_complete\_var} & 66.67\\
\hline
\end{tabular}
\caption{\label{tab:n-prop-pct-miss-complete}Single number summaries of
missingness and completeness of the airquality dataset.  The functions
follow consistent naming, making them easy to remember, and their use
clear.}
\end{table}

\hypertarget{sum-tab-missings}{%
\subsection{Summaries and tabulations of missing data}\label{sum-tab-missings}}

Presenting the number and percent of missing values for each variable, or
case, provides a summary usable in data handling, or in models to inform
imputation.  For example, potentially dropping variables, or deciding to
include others in an imputation model.  Another useful approach is to
tabulate the frequency of missing values for each variable or case; that is,
the number of times there are zero missings, one missing, two, and so on.
These summaries and tabulations are shown for variables in
Tables~\ref{tab:miss-var-summary} and~\ref{tab:miss-var-table}, and
implemented with \code{miss_var_summary()} and \code{miss_var_table()}.
Case-wise (row-wise) summaries and tabulations are implemented with
\code{miss_case_summary()} and \code{miss_case_table()}.

These summaries order rows by the number of missings (\code{n_miss()}), to
show the most missings at the top.  The number of missings across a
repeating span, or finding ``runs'' or ``streaks'' of missings in given
variables can also be useful for identifying missingness patterns, and are
implemented with \code{miss_var_span()}, and \code{miss_var_run()}.

\begin{table}[t!]
\centering
\begin{tabular}[t]{lrr}
\hline
\code{Variable} & \code{n\_miss} & \code{pct\_miss}\\
\hline
Ozone & 37 & 24.2\\
Solar.R & 7 & 4.6\\
Wind & 0 & 0.0\\
Temp & 0 & 0.0\\
Month & 0 & 0.0\\
Day & 0 & 0.0\\
\hline
\end{tabular}
\caption{\label{tab:miss-var-summary}\code{miss\char`_var\char`_summary(airquality)}
provides the number and percent of missings in each variable in airquality.
Only ozone and solar radiation have missing values.}
\end{table}

\begin{table}[t!]
\centering
\begin{tabular}[t]{rrr}
\hline
\code{n\_miss\_in\_var} & \code{n\_vars} & \code{pct\_vars}\\
\hline
0 & 4 & 66.7\\
7 & 1 & 16.7\\
37 & 1 & 16.7\\
\hline
\end{tabular}
\caption{\label{tab:miss-var-table}The output of
\code{miss\char`_var\char`_table(airquality)}, tabulating the amount of
missing data in each variable in airquality.  This shows the number of
variables with 0, 7, and 37 missings, and the percentage of variables with
those amounts of missingness.  There are few missingness patterns.}
\end{table}

\hypertarget{num-sum-w-group}{%
\subsection{Combining numerical summaries with grouping operations}\label{num-sum-w-group}}

It is useful to explore summaries and tabulations within groups of a
dataset.  \pkg{naniar} works with \pkg{dplyr}'s \code{group_by()} operator
to produce grouped summaries, which work well with the ``pipe'' operator.
Table~\ref{tab:group-miss-var-summary} below shows an example of missing
data summaries for airquality, grouped by month.

\begin{table}[t!]
\centering
\begin{tabular}[t]{rlrr}
\hline
\code{Month} & \code{Variable} & \code{n\_miss} & \code{pct\_miss}\\
\hline
5 & Ozone & 5 & 16.1\\
5 & Solar.R & 4 & 12.9\\
5 & Wind & 0 & 0.0\\
5 & Temp & 0 & 0.0\\
5 & Day & 0 & 0.0\\
6 & Ozone & 21 & 70.0\\
6 & Solar.R & 0 & 0.0\\
6 & Wind & 0 & 0.0\\
6 & Temp & 0 & 0.0\\
6 & Day & 0 & 0.0\\
\hline
\end{tabular}
\caption{\label{tab:group-miss-var-summary}Output of \code{airquality \%>\%
group\char`_by(Month) \%>\% miss\char`_var\char`_summary()} provides a
grouped summary of the missingness in each variable, for each month of the
airquality dataset.  Only the first 10 rows are shown.  There are more ozone
missings in June than May.}
\end{table}

\hypertarget{case-study}{%
\section{Application}\label{case-study}}

This section shows how the methods described so far are used together in a
data analysis workflow.  We analyse a case study of housing data for the
city of Melbourne from January 28, 2016 to March 17, 2018.  The data was
compiled by scraping weekly property clearance data
\citep{Kaggle-2018-data}.  There are $27,247$ properties, and 21~variables in
the dataset.  The variables include real estate type (town house, unit,
house), suburb, selling method, number of rooms, price, real estate agent,
sale date, and distance from the central business district.

The goal in analyzing this data is to accurately predict Melbourne housing
prices.  The data contains many missing values.  As a precursor to building
a predictive model, this analysis focuses on understanding the patterns of
missingness.

\hypertarget{case-study-explore-pattern}{%
\subsection{Exploring patterns of missingness}\label{case-study-explore-pattern}}

Figure~\ref{fig:housing-miss-case-var}A shows 9 variables with missing
values.  The most missings are in ``building area'', followed by ``year
built'', and ``land size'', with similar amounts of missingness in ``Car'',
``bathroom'', ``bedroom2'', ``longitude'', and ``latitude''.
Figure~\ref{fig:housing-miss-case-var}B reveals there are up to 50\% missing
values in cases, and the majority of cases have more than 5\% values
missing.  The variables ``building area'' and ``year built'' have more than
50\% missing data, and so could perhaps be omitted from subsequent analysis,
as imputed values are likely to be spurious.  Three missingness clusters are
revealed by visualizing missingness in the whole dataset, clustering and
arranging the rows and columns of the data
(Figure~\ref{fig:applic-vis-miss}).

\begin{figure}[t!]
\centering
\includegraphics[width=1\linewidth]{housing-miss-case-var-1}
\caption[The amount of missings in variables (A) and cases (B) for Melbourne
housing data]{The amount of missings in variables (A) and cases (B) for
Melbourne housing data.  (A) Build area and year built have more than 50\%
missing, and car, bathroom, bedroom2 and longitude and latitude have about
25\% missings.  (B) Cases are missing $5 - 50\%$ of values.  The majority of
missingness is in selected cases and
variables.}\label{fig:housing-miss-case-var}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics[width=0.85\linewidth]{applic-vis-miss-1}
\caption[Heatmap of clustered missingness for housing data reveals
structured missingness]{Heatmap of clustered missingness for housing data
reveals structured missingness.  Three groups of missingness are apparent.
At the top: building area to longitude; the middle: building area and year
built; the end: building area, year built, and
landsize.}\label{fig:applic-vis-miss}
\end{figure}

Figure~\ref{fig:housing-upset} shows missingness patterns with an upset
plot \citep{Conway2017}, displaying 8 intersecting sets of missing
variables.  Two patterns stand out: two, and five variables missing,
providing further evidence of the missingness patterns seen in
Figures~\ref{fig:housing-miss-case-var} and~\ref{fig:housing-upset}.

\begin{figure}[t!]
\centering
\includegraphics[width=1\linewidth]{housing-upset-1}
\caption[An upset plot of 8 sets of missingness in the housing data]{An
upset plot of 8 sets of missingness in the housing data.  Missingness for
each variable is shown on the bottom left.  Connected dots show
co-occurrences of missings in variables.  Two missingness patterns are clear,
year built and building area, and latitude through to building
area.}\label{fig:housing-upset}
\end{figure}

Tabulating the number of missings in variables in
Table~\ref{tab:housing-miss-var-case-table} (left) shows three groups of
missingness.  Tabulating missings in cases
(Table~\ref{tab:housing-miss-var-case-table}, right) shows six patterns of
missingness.  These overview plots lead to the removal of two variables from
with more than 50\% missingness from analysis: ``building area'' and ``year
built''.

\begin{table}
\centering
\begin{tabular}[t]{rrr}
\hline
\code{n\_miss\_in\_var} & \code{n\_vars} & \code{pct\_vars}\\
\hline
0 & 10 & 47.6\\
1 & 2 & 9.5\\
3 & 1 & 4.8\\
6254 & 2 & 9.5\\
6441 & 1 & 4.8\\
6447 & 1 & 4.8\\
6824 & 1 & 4.8\\
9265 & 1 & 4.8\\
15163 & 1 & 4.8\\
16591 & 1 & 4.8\\
\hline
\end{tabular}
\centering
\begin{tabular}[t]{rrr}
\hline
\code{n\_miss\_in\_case} & \code{n\_cases} & \code{pct\_cases}\\
\hline
0 & 8887 & 32.6\\
1 & 3237 & 11.9\\
2 & 7231 & 26.5\\
3 & 1370 & 5.0\\
4 & 79 & 0.3\\
5 & 8 & 0.0\\
6 & 203 & 0.7\\
8 & 6229 & 22.9\\
9 & 2 & 0.0\\
11 & 1 & 0.0\\
\hline
\end{tabular}
\caption{\label{tab:housing-miss-var-case-table}Summary tables to help
understand missingness patters.  Output of \code{miss\_var\_table(housing)}
(left), tabulating missingness for variables, and output of
\code{miss\_case\_table(housing)}.  There are 13 variables with $0-3$
missings, 6 variables have $6,000-10,000$ missings, 2 variables have $15,000 -
17,000$ missings.  About 30\% of cases have no missings, 45\% of cases have $1
- 6$ missings, and about 23\% of cases have 8 or more missings.  There are
different patterns of missingness in variables and cases, but they can be
broken down into smaller groups.}
\end{table}

\hypertarget{case-study-explore-for-imp}{%
\subsection{Exploring missingness patterns for imputation}\label{case-study-explore-for-imp}}

Using information from Section~\ref{case-study-explore-pattern}, the following
variables are explored for features predicting missingness: ``latitude'',
``longitude'', ``bedroom2'', ``bathroom'', ``car'', and ``land size''.

Missingness structure is explored by clustering the missing values into
groups.  Then, a classification and regression trees (CART) model is applied
to predict these missingness clusters using the remaining variables
\citep{Tierney2015, Barnett2017}.  Two clusters of missingness are
identified and predicted using all variables in the dataset with the CART
package \pkg{rpart} \citep{rpart}, and plotted using the \pkg{rpart.plot}
package \citep{rpart-plot}.  Importance scores reveal the following
variables as most important in predicting missingness: ``rooms'', ``price'',
``suburb'', ``council area'', ``distance'', and ``region name''.  These
variables are important for predicting missingness, so are included in the
imputation model.

\begin{figure}[t!]
\centering
\includegraphics[width=0.9\linewidth]{rpart-plot-1}
\caption[Decision tree output predicting missingness clusters]{Decision tree
output predicting missingness clusters.  Type of house, year quarter, and
year were important for predicting missingness cluster.  The cluster with
the most missingness was for quarters 1 and 4, for 2017 and 2018.  Type of
house, year, and year quarter are important features related to missingness
structure.}\label{fig:rpart-plot}
\end{figure}

\hypertarget{case-study-imp-diagnosis}{%
\subsection{Imputation and diagnostics}\label{case-study-imp-diagnosis}}

\pkg{simputation} is used to implement two imputations: simple linear
regression and $K$ nearest neighbors.  Values are imputed stepwise in
ascending order of missingness.  The track missings pattern is applied
(described in Section~\ref{verbs}), to assess imputed values.  Imputed datasets are
compared on their performance in a model predicting log house price for 4~variables
(Figure~\ref{fig:imputed-by-model}).  Compared to $k$-nearest neighbors (KNN) imputed
values, the linear model imputed values closer to the mean.

\begin{figure}[t!]
\centering
\includegraphics[width=0.95\linewidth]{imputed-by-model-1}
\caption[Boxplots of complete case data, and data imputed with KNN or linear
model for different variables]{Boxplots of complete case data, and data
imputed with KNN or linear model for different variables.  (A) number of
bedrooms, (B) number of bathrooms, (C) number of carspots, and (D) landsize
(on a log10 scale).  KNN had similar results to complete case, and linear
model had a lower median for cars and fewer extreme values for
bedrooms.}\label{fig:imputed-by-model}
\end{figure}

\hypertarget{case-study-assess-model}{%
\subsection{Assessing model predictions}\label{case-study-assess-model}}

Coefficients of the linear model of log price vary for room for different
imputed datasets (Figure~\ref{fig:tidy-coefs}).  Notably, complete cases
underestimate the impact of room on log price.  A partial residual plot
(Figure~\ref{fig:partial-resid}) shows there is not much variation amongst
the models from the differently imputed datasets.

\begin{figure}[t!]
\centering
\includegraphics[width=1\linewidth]{tidy-coefs-1}
\caption[The coefficient estimate for the number of rooms varies according
to the imputed dataset]{The coefficient estimate for the number of rooms
varies according to the imputed dataset.  Complete case dataset produced
lower coefficients, compared to imputed datasets.}\label{fig:tidy-coefs}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{partial-resid-1}
\caption[Partial residual plot for each data set, complete cases (cc), and
imputed with KNN (knn) or linear model (lm)]{Partial residual plot for each
data set, complete cases (cc), and imputed with KNN (knn) or linear model
(lm).  These are plotted as hex bins, colored according to the number of
points in a given hexagon.  Brighter colors mean more points.  Compared to
complete cases, imputed data has more points clustered around
zero.}\label{fig:partial-resid}
\end{figure}

\hypertarget{case-study-summary}{%
\subsection{Summary}\label{case-study-summary}}

The \pkg{naniar} and \pkg{visdat} packages implement the methods discussed
in the paper, building on existing tidy tools and strike a compromise
between automation and control, making analysis efficient, readable, but not
overly complex.  Each tool has clear intent and effects -- summarizing,
plotting or generating data or augmenting data in some way.  This not only
reduces repetition and typing in an analysis, but most importantly, allows
for clear expression of intent, making exploration of missing values fluent.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This paper has described new methods for exploring, visualizing, and
imputing missing data.  The work was motivated by recent developments of
tidy data, and extends them for better missing value handling.  The methods
have standard outputs, function arguments, and behavior.  This provides
consistent workflows centered around data analysis that integrate well with
existing imputation methodology, visualization, and modelling.

The \emph{nabular} data structures discussed in the paper are simple by
design, to promote flexibility.  They could be used to create different
visualizations than were shown in the paper.  The analyst can use the data
structures to decide on appropriate visualization for their problem.  The
data structures could also be used to support interactive graphics, in the
manner of \proglang{MANET} and \proglang{ggobi}.  Linking the plots (via
linked brushing) could facilitate exploration of missingness, and could be
implemented with \pkg{plotly} \citep{plotly} for added interactivity.
Animating between different sets of imputed values could also be possible
with packages like \pkg{gganimate} \citep{gganimate}.

Other data structures such as spatial data, time series, networks, and
longitudinal data would be supported by the inherently tabular,
\emph{nabular} data, if they are first structured as wide tidy format.
Large data may need special handling, and additional features like efficient
storage of purely imputed values and lazy evaluation.  Special missing value
codes could be improved by creating special classes, or expanding low level
representation of NA at the source code level.  Similarly, list
columns, present a challenge for analysis as their expansion into vectors in
a dataframe is not always possible as they have different lengths, but
should still be flagged for further analysis.

The methodology described in this paper can be used in conjunction with
other approaches to understand multivariate missingness dependencies (e.g.,~decision
trees \citep{Tierney2015}, latent group analysis
\citep{Barnett2017}, and PCA \citep{FactoMineR}).  Evaluating imputed values
using a testing framework like \citep{vanBuuren2018} is also supported.

The approach meshes with the dynamic nature of data analysis, allowing the
analyst to go from raw data to model data in a fluid workflow.

\hypertarget{acknowledgements}{%
\section*{Acknowledgments}\label{acknowledgements}}

The authors would like to thank Miles McBain, for his key contributions and
discussions on the \pkg{naniar} package, in particular for helping implement
\code{geom_miss_point()}, and for his feedback on ideas, implementations,
and names.  We also thank Colin Fay for his contributions to the
\pkg{naniar} package, in particular for his assistance with the
\code{replace_with_na()} functions.  We also thank Earo Wang and Mitchell
O'Hara-Wild for the many useful discussions on missing data and package
development, and for their assistance with creating elegant approaches that
take advantage of the tidy syntax.  We would also like to thank those who
contributed pull requests and discussions on the \pkg{naniar} package, in
particular Jim Hester and Romain Fran{\c{c}}ois for improving the speed of key
functions, Ross Gayler for discussion on special missing values, and Luke
Smith for helping \pkg{naniar} be more compliant with \pkg{ggplot2}.  We
would also like to thank Amelia McNamara for discussions on the paper.

\bibliography{ref}

\end{document}